---
title: "Robbery in Central San Francisco during the Summer of 2014"
author: "Dan Kirlin"
date: "December 29, 2015"
output: pdf_document
---

# Objective 

In this project, we will be examining Crime in San Francisco from the Summer of 2014 using the `R` programming language and `ggplot2` package. In our investigation we will show several trends for San Francisco during this time period: 

1. Theft is the most common crime type.
2. The Central District, although not a top 3 district for crime overall, is top 2 in theft and/or robberies.
3. There is no apparent trend towards Days for this type of crime in the Central District.

# Paper

Let's begin by loading the ggplot2 package, dplyr (data manipulation), and the provided data.

```{r, message=FALSE}
require(ggplot2);require(dplyr);
sanfrancisco <- read.csv("sanfrancisco_incidents_summer_2014.csv")
```

For our first graph, let's examine how Crime is dispersed throughout the various districts. 
```{r}
plot_neighborhood <- ggplot(sanfrancisco, aes(x = PdDistrict))
plot_neighborhood <- plot_neighborhood + geom_bar()
plot_neighborhood <- plot_neighborhood + labs(y = "Frequency", x = "District")
plot_neighborhood <- plot_neighborhood + theme(axis.text.x=element_text(angle=45,vjust=.6))
plot_neighborhood <- plot_neighborhood + ggtitle("Frequency of Crime in San Francisco by District")
plot_neighborhood
```
Here we see that the Southern District has nearly double that of the next highest (Central, Mission, and Northern).

Let's now examine this from a different angle - what type of crimes are taking place in each district?
```{r}
plot <- ggplot(sanfrancisco, aes(x = Category, fill=PdDistrict))
plot <- plot + theme(axis.text.x=element_text(angle=45,vjust=.6))
plot <- plot + geom_bar()
plot <- plot + xlab("Frequency")
plot <- plot + ylab("Crime Type")
plot <- plot + ggtitle("San Francisco Crime Type by District")
plot <- plot + scale_fill_discrete(name = "District")
plot
```
The most common crime by nearly double is Larceny/Theft. This appears to happen most frequently in the Central and Southern district.

# Robbery Examination
From a purely Robbery/Theft perspective we see that central is only second to that of southern. Which when compares to only 4th overall for crime of the districts, is a significant jump.
```{r}
sanfrancisco_theft <- filter(sanfrancisco, Category == 'ROBBERY' | Category == 'LARCENY/THEFT')
plot_theft <- ggplot(sanfrancisco_theft, aes(x=PdDistrict))
plot_theft <- plot_theft + geom_bar()
plot_theft <- plot_theft + ggtitle("Robbery and Theft in San Francisco")
plot_theft <- plot_theft + labs(x = "District", y = "Frequency")
plot_theft <- plot_theft + theme(axis.text.x=element_text(angle=45,vjust=.6))
plot_theft
```

# Central District Examination
If we narrow our scope to only the Central District exclusively, we can see several trends
```{r}
sanfrancisco_center <- filter(sanfrancisco, PdDistrict=="CENTRAL")
plot_incidents <- ggplot(sanfrancisco_center, aes(x=Category))
plot_incidents <- plot_incidents + geom_bar()
plot_incidents <- plot_incidents + theme(axis.text.x=element_text(angle=45,vjust=.6))
plot_incidents <- plot_incidents + labs(x = "Crime Type", y = "Frequency")
plot_incidents <- plot_incidents + ggtitle("Crime Type in Central San Francisco")
plot_incidents
```
For one - Larceny and Theft is even more strongly represented in this district (nearly quadruple the next crime type of "non-criminal"). 

Let's now attempt to see if there are any trends related to Day of the Week.
```{r}
sanfrancisco_central_theft <- filter(sanfrancisco_center, Category == 'ROBBERY' | Category == 'LARCENY/THEFT')
plot_center_theft <- ggplot(sanfrancisco_central_theft, aes(x=DayOfWeek))
plot_center_theft <- plot_center_theft + geom_bar()
plot_center_theft <- plot_center_theft + labs(x="Day of the Week", y="Frequency")
plot_center_theft
```
Unfortunately - from the following graph we can see no trend in robbery to days of the week. One would asssume that Weekend nights would be a hot bed for this kind of activity however, data never lies. (A good point for a future project would be to visualize this data as a time series.)

# Conclusion
In conclusion - we can see that Robbery is largely represented in the Central District as a percentage of it's crime, is the second overall for this type of crime among all districts and appears to have no trends with days of the week.
